Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}

```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}

```

***

### Overplotting
Notes:

```{r Overplotting}
library(ggplot2)
data <- read.csv("pseudo_facebook.tsv", sep="\t")
ggplot(aes(x = age, y = friend_count), data = data) + 
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) + 
  xlim(13, 90) + 
  coord_trans(y = "sqrt")
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}

```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friendships_initiated), data = data) + 
  geom_point(alpha = 0.1, position = position_jitter(h = 0)) + 
  xlim(13, 110) + 
  coord_trans(y = "sqrt")

```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library(dplyr)
fc_by_age<- data %>%
  group_by(age) %>%
  summarise(fc_mean = mean(friend_count), 
            fc_median = median(friend_count), 
            n = n()) %>%
  arrange(age)

head(fc_by_age)
```

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x = age, y = fc_mean), data = fc_by_age) +
  geom_line()
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = data) + 
  geom_point(alpha = 1/20, position = position_jitter(h = 0), color = "orange") + 
  coord_cartesian(xlim = c(13, 70), ylim = c(0, 1000)) + 
  geom_line(stat = "summary", fun.y = "mean") +
  geom_line(stat = "summary", fun.y = "median", color = "blue") + 
  geom_line(stat = "summary", fun.y = "quantile", fun.args = list(probs = .1), color = "blue", linetype = 2) +  
  geom_line(stat = "summary", fun.y = "quantile", fun.args = list(probs = .9), color = "blue", linetype = 2)
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
?cor.test
cor.test(data$age, data$friend_count, method = "pearson")
with(data, cor.test(age, friend_count, method = "pearson"))
with(subset(data, age <= 70), cor.test(age, friend_count))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(                 , cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(x = www_likes_received, y = likes_received), data = data) + 
  geom_point(alpha = 0.25) + 
  xlim(0, quantile(data$www_likes_received, 0.95)) +
  ylim(0, quantile(data$likes_received, 0.95)) +
  geom_smooth(method = "lm", color = "red")
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
with(data, cor.test(x = www_likes_received, y = likes_received))
```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
library(alr3)
data(Mitchell)
?Mitchell
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
with(Mitchell, cor.test(x = Month, y = Temp))
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point() +
  scale_x_continuous(breaks = seq(0, 200, 12))
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

dcor.ttest(x, y) in energy package can catch this pattern better
***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
data$age_with_months <- with(data, age + (1 - dob_month / 12))
```

***

### Age with Months Means

```{r Age with Months Means}
fc_by_age_month<- data %>%
  group_by(age_with_months) %>%
  summarise(fc_mean = mean(friend_count), 
            fc_median = median(friend_count), 
            n = n()) %>%
  arrange(age_with_months)

head(fc_by_age_month)
```

Programming Assignment
```{r Programming Assignment}
ggplot(aes(x = age_with_months, y = fc_mean), data = subset(fc_by_age_month, age_with_months < 71)) + 
  geom_line()
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(aes(x = age, y = fc_mean), data = subset(fc_by_age, age < 71)) +
  geom_line() +
  geom_smooth()
  
p2 <- ggplot(aes(x = age_with_months, y = fc_mean), data = subset(fc_by_age_month, age_with_months < 71)) + 
  geom_line() + 
  geom_smooth()

p3 <- ggplot(aes(x = round(age / 5) * 5, y = friend_count), data = subset(data, age < 71)) + 
  geom_line(stat = "summary", fun.y = mean) 

library(gridExtra)
grid.arrange(p1, p2, p3, ncol = 1)
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

