Lesson 4
========================================================
### data uploading
```{r diamonds}
library(ggplot2)
data("diamonds")
?diamonds
dim(diamonds)
sapply(diamonds, class)
levels(diamonds$color)
```
***

### price histogram center
```{r price hist}
mean(diamonds$price)
median(diamonds$price)
```

***
### dimond counts
```{r price hist}
sum(diamonds$price < 500)
sum(diamonds$price < 250)
sum(diamonds$price >=  15000)
```

***
# Explore the largest peak in the
# price histogram you created earlier.

# Try limiting the x-axis, altering the bin width,
# and setting different breaks on the x-axis.

# There won’t be a solution video for this
# question so go to the discussions to
# share your thoughts and discover
# what other people find.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).
```{r price hist}
ggplot(aes(diamonds$price), data=diamonds) +
    geom_histogram(binwidth = 1) +
    scale_x_continuous(limits = c(0, 1000))
```

***
# Break out the histogram of diamond prices by cut.

# You should have five histograms in separate
# panels on your resulting plot.
```{r price hist}
ggplot(aes(diamonds$price), data=diamonds) +
    geom_histogram(binwidth = 30) +
    facet_wrap(~cut)
```
# Break out the histogram of diamond prices by cut.

# You should have five histograms in separate
# panels on your resulting plot.
```{r price hist}
ggplot(aes(price), data=diamonds) +
    geom_histogram() +
    facet_wrap(~cut)                                                 
```
# price by cut questions

```{r price hist}
by(diamonds$price, diamonds$cut, summary)
```
# The 'Fair' and 'Good' diamonds appear to have 
# different distributions compared to the better
# cut diamonds. They seem somewhat uniform
# on the left with long tails on the right.

# Let's look in to this more.

# Look up the documentation for facet_wrap in R Studio.
# Then, scroll back up and add a parameter to facet_wrap so that
# the y-axis in the histograms is not fixed. You want the y-axis to
# be different for each histogram.in separate
# panels on your resulting plot.
```{r price hist}
?facet_wrap
ggplot(aes(price), data=diamonds) +
    geom_histogram(binwidth = 30) +
    facet_wrap(~cut, scales = "free_y")
```
# Create a histogram of price per carat
# and facet it by cut. You can make adjustments
# to the code from the previous exercise to get
# started.

# Adjust the bin width and transform the scale
# of the x-axis using log10.
```{r price hist}
ggplot(aes(price / carat), data=diamonds) +
    geom_histogram() +
    facet_wrap(~cut) +
    scale_x_log10()
```
# Investigate the price of diamonds using box plots,
# numerical summaries, and one of the following categorical
# variables: cut, clarity, or color.

# There won’t be a solution video for this
# exercise so go to the discussion thread for either
# BOXPLOTS BY CLARITY, BOXPLOT BY COLOR, or BOXPLOTS BY CUT
# to share you thoughts and to
# see what other people found.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Copy and paste all of the code that you used for
# your investigation, and submit it when you are ready.
# Price by clarity
```{r price by clarity}
by(diamonds$price, diamonds$clarity, summary)
ggplot(aes(x=clarity, y=price), data=diamonds) +
    geom_boxplot() +
    coord_cartesian(ylim = c(0, 7000))
```
# price by color
```{r price by color}
by(diamonds$price, diamonds$color, summary)
ggplot(aes(x=color, y=price), data=diamonds) +
    geom_boxplot() +
    coord_cartesian(ylim = c(0, 10000))
```
# Investigate the price per carat of diamonds across
# the different colors of diamonds using boxplots.

# Go to the discussions to
# share your thoughts and to discover
# what other people found.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Copy and paste all of the code that you used for
# your investigation, and submit it when you are ready.
```{r price hist}
ggplot(aes(x = color, y = price / carat), data=diamonds) +
    geom_boxplot() +
    coord_cartesian(ylim = c(0, 7000))
```
# Frequency polygon of carats

```{r price hist}
ggplot(aes(carat), data=diamonds) +
    geom_freqpoly(bintwidth=50) +
    scale_y_continuous(limit = c(0, 2000)) +
    scale_x_continuous(limit = c(0, 2), breaks = seq(0, 2, 0.1))
```
# The Gapminder website contains over 500 data sets with information about
# the world's population. Your task is to download a data set of your choice
# and create 2-5 plots that make use of the techniques from Lesson 3.

# You might use a simple histogram, a boxplot split over a categorical variable,
# or a frequency polygon. The choice is yours!

# You can find a link to the Gapminder website in the Instructor Notes.

# Once you've completed your investigation, create a post in the discussions that includes:
#       1. any questions you answered, your observations, and summary statistics
#       2. snippets of code that created the plots
#       3. links to the images of your plots
```{r price hist}
library(xlsx)
library(tidyr)
library(dplyr)
TB_raw <- read.xlsx("indicator_all tb deaths per 100000.xlsx", sheetName="Data")
TB_raw[, 2:19][is.na(TB_raw[, 2:19])] <- 0
colnames(TB_raw)[1] <- 'country'
head(TB_raw)
TB_data <- gather(TB_raw, 'year', 'deaths_per_100k', 2:19)
head(TB_data)
ggplot(aes(x = year, y = deaths), data = subset(TB_data, !is.na(deaths))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(0, 100)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
# plot 2 about tb deaths

```{r price hist}
TB_total_raw <- summarise_each(TB_raw, funs(sum), 2:19)
TB_total <- gather(TB_total_raw, 'year', 'total_deaths_per_100k', 1:18)
ggplot(aes(x=year, y=total_deaths_per_100k, group=1), data=TB_total) +
  geom_line(stat="identity", color="green") +
  geom_point(size=3, color="blue") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
# Your task is to investigate the distribution of your friends'
# birth months and days.

# Here some questions you could answer, and we hope you think of others.

# **********************************************************************

# How many people share your birthday? Do you know them?
# (Reserve time with them or save money to buy them a gift!)
0
# Which month contains the most number of birthdays?
May

# How many birthdays are in each month?

# Which day of the year has the most number of birthdays?

# Do you have at least 365 friends that have birthdays on everyday
# of the year?
No
```{r price hist}
fb_bd <- read.csv("fb_birthdays.csv", sep='\t')
fb_bd <- fb_bd[,1:2]
colnames(fb_bd) <- c("name", "date")
formatted_date <- strptime(fb_bd$date, "%d.%m.%Y")
fb_bd$day <- format(formatted_date, "%d")
fb_bd$month <- format(formatted_date, "%m")
fb_bd$tmp <- rep(1, 96)
head(fb_bd)
by(fb_bd$day, fb_bd$month, length)
by(fb_bd$month, fb_bd$day, length)
subset(fb_bd, month == "07")
```

```